How Slow do my Transient Typechecks Go?

### 2019-08-17 graphs/preamble.tex
Added Section 4

% Available ACM colours: ACMBlue, ACMYellow, ACMOrange, ACMRed, ACMLightBlue, ACMGreen, ACMPurple and ACMDarkBlue
\colorlet{GraphNormal}{ACMRed} % Normal colour when I don't care about having multiple colours

@@

\pgfplotsset{%
% WARNING: do not put empty lines here, and make sure each = ... is terminated with a comma!
	every axis/.append style = {% options for all axes/graphs
		cycle list={}, tick align=center, xtick pos = bottom, ymin=0, axis on top,%
		axis y line*=left, try min ticks=5,%
		every tick label/.append style={font=\tiny},%
		every axis label/.append style={font=\tiny},%
	},%
	%
	Full/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=2pt}, ymax={#1}, height=3.2cm, width=5.9cm,%
		every axis plot/.style={mark size=1pt, draw=GraphNormal, fill=GraphNormal!40}},%
	Right Axis/.style = { % For right axes
		axis y line*=right, axis x line=none, xmin=0, xmax=100},% 
	%
	Scatter/.style = { % For the scatter plots
		xticklabel={$\pgfmathprintnumber{\tick}$\%},%
		every axis plot/.append style={only marks}, mark=o},%
	%
	% Variants of the above that contain axis labels
	Scatter Left Label/.style   = {ylabel={Average Time (ms)}},%
	Scatter Bottom Label/.style = {xlabel={Type Annotations}},%
	Scatter Right Label/.style  = {ylabel={Relative to 0\% Typed}},%
	%
	%
	Column/.style = {% For the column graphs
		enlarge x limits=false, xtick distance={#1}, const plot,%
		every axis plot/.append style={no markers}},%
	%
	% Variants of the above that contain axis labels
	Column Left Label/.style   = {ylabel={Average Time (ms)}},%
	Column Bottom Label/.style = {xlabel={Type Annotation Index}},%
	Column Right Label/.style  = {ylabel={Relative to Untyped}},%
	%
	Pattern/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=1.5pt}, ymax={#1}, height=3.2cm, width=4.4cm, %
		every axis plot/.style={mark size=0.75pt, draw=GraphBase, fill=GraphBase!40}},%
	%
	Pattern A/.style = {draw=GraphA, fill=GraphA!40, mark=*}, %
	Pattern B/.style = {draw=GraphB, fill=GraphB!40, mark=*}, %
	Pattern AB/.style = {draw=GraphAB, fill=GraphAB!40, mark=*}, %
	%Pattern Left/.style = {Left Axis={#1}, height=3.2cm, width=7.9cm}, % Size for pattern graphs
	%Full Scatter Left/.style = {Scatter, Full Left={#1},
	%	enlarge x limits={abs=2pt}, every axis plot/.append style={mark size={1pt}}},%
	%Full Scatter Right/.style       = {Full Scatter Left={#1},        Right Axis},%
	%sleft = "Pattern Scatter Left"
	%cleft = "Pattern Column Left"
	%sright = "Pattern Scatter Right"
	%cright = "Pattern Column Right"
	%Pattern Scatter A, Pattern Scatter B
}

>>>>

% Available ACM colours: ACMBlue, ACMYellow, ACMOrange, ACMRed, ACMLightBlue, ACMGreen, ACMPurple and ACMDarkBlue
\colorlet{GraphNormal}{ACMOrange} % Normal colour when I don't care about having multiple colours

@@

\pgfplotsset{%
% WARNING: do not put empty lines here, and make sure each = ... is terminated with a comma!
	every axis/.append style = {% options for all axes/graphs
		cycle list={}, tick align=center, xtick pos = bottom, ymin=0, axis on top,%
		axis y line*=left, try min ticks=5,%
		every tick label/.append style={font=\tiny},%
		every axis label/.append style={font=\tiny},%
	},%
	%
	Full/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=1.5pt}, ymax={#1}, height=3.2cm, width=5.9cm,%
		every axis plot/.style={very thin, mark size=0.75pt, draw=GraphNormal, fill=GraphNormal!40}},%
	Right Axis/.style = { % For right axes
		axis y line*=right, axis x line=none, xmin=0, xmax=100},% 
	%
	Scatter/.style = { % For the scatter plots
		xticklabel={$\pgfmathprintnumber{\tick}$\%},%
		every axis plot/.append style={only marks}, mark=o},%
	%
	% Variants of the above that contain axis labels
	Scatter Left Label/.style   = {ylabel={Average Time (ms)}},%
	Scatter Bottom Label/.style = {xlabel={Proportion of Type Annotations}},%
	Scatter Right Label/.style  = {ylabel={Relative to 0\% Typed}},%
	%
	%
	Column/.style = {% For the column graphs
		enlarge x limits=false, xtick distance={#1}, const plot,%
		every axis plot/.append style={no markers}},%
	%
	% Variants of the above that contain axis labels
	Column Left Label/.style   = {ylabel={Average Time (ms)}},%
	Column Bottom Label/.style = {xlabel={Type Annotation Index}},%
	Column Right Label/.style  = {ylabel={Relative to Untyped}},%
	%
	Pattern/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=1pt}, ymax={#1}, height=3.2cm, width=4.4cm, % 4.4
		every axis plot/.style={very thin, mark size=0.5pt, draw=GraphBase, fill=GraphBase!40}},%
	%
	Pattern A/.style = {draw=GraphA, fill=GraphA!40, mark=*}, %
	Pattern B/.style = {draw=GraphB, fill=GraphB!40, mark=*}, %
	Pattern AB/.style = {draw=GraphAB, fill=GraphAB!40, mark=*}, %
}

### 2019-08-17 paper.tex
Added Section 4

Though though outliers, caused by \eg garbage collection, remain visible in the plots for each individual benchmark, the largest 95\% confidence interval we obtained for the arithmetic mean of any of the results for any of our experiments was $\pm3.53\%$ (for the PyStone benchmark).

All experiments were executed on a Windows 10 machine
running Fedora 29 within Windows Subsystems for Linux (version 1). The machine has

one Intel Core i7-6800K @3.40GHz, with 6 cores, for a total of 12 hyperthreads. We used ReBench 1.0 \citep{ReBench:2018}, Java 1.8.0\_191 Graal 0.43.
Benchmarks were executed one by one to avoid interference between them.
The analysis of the results and plots where generated using Python 3.7.3 and PGFPLOTS 1.16.
To enable reproductions, the scripts we used to generate and run our experiments, including the source code for all the configurations tested, is available online.\footnote{
\url{https://gitlab.ecs.vuw.ac.nz/isaac/Moth-Takikawa}}

@@

give overall results - one page of monochromatic scatterplots one for
each benchmark.

need to explKn how to read scatterplots by paraphrasing
\cite{vitousek-transient-arXive-2019}

core results - i.e.\ looking at one scatter plot per annotation - the original version
   - is there a good benchmark that is small enough to fit on one
   page. and shows interesting results?

\begin{figure*}
	\input{graphs/full-scatter}
	\caption{The graphs!}
\end{figure*}

\section{Individual Annotations}
\label{s-individual}

>>>>

Though outliers, caused by \eg garbage collection, remain visible in the plots for each individual configuration, the largest 95\% confidence interval we obtained (over the mean time after warmup) for any of experiments was $\pm3.53\%$ (for the PyStone benchmark).

All experiments were executed on a Windows 10 machine
running Fedora 30 within Windows Subsystems for Linux (version 1). The machine has an Intel Core i7-6800K @3.40GHz, with 6 cores, for a total of 12 hyperthreads. We used ReBench 1.0 \citep{ReBench:2018}, Java 1.8.0\_191 Graal 0.43.
Benchmarks were executed one by one to avoid interference between them.
The analysis of the results and plots where generated using Python 3.7.3 and PGFPLOTS 1.16.
To enable reproductions, the scripts we used to generate and run our experiments, including the source code for all the configurations tested, is available online.\footnote{
\url{https://gitlab.ecs.vuw.ac.nz/isaac/Moth-Takikawa}}

@@

\begin{figure*}
	\input{graphs/full-scatter}
	\caption{Graphs of (at most) 102 configurations in the typing latices for each benchmark. Time is measured as the mean of the 351\textsuperscript{st} to the 1,000\textsuperscript{th} benchmark iteration under a single invocation of Moth (lower is better).}
	\label{f:full}
\end{figure*}
The results for the performance of our sample of typing lattice configurations are presented in figure~\ref{f:full}. Following \cite{vitousek-transient-arXive-2019}, the points on each graph show the average execution if each individual configuration. The x-axis represents the proportion of type annotations for each configuration, with the left and right most points showing the times for the fully untyped and typed configurations (respective). The execution time (in milliseconds) is shown on the left y-axes, and time relative to the fully untyped configuration is shown on the right y-axes.

Most of the graphs are mostly horizontal lines: indicating that the performance effect of including type annotations is negligible. However the plots for CD and Richards show a roughly linear increase: i.e. adding more type annotations to the program increases performance linearly. The plots for Go, Permute, DeltaBlue, and Storage even show \emph{decreases}: i.e. adding more type annotations can actually \emph{increase} performance. Of particular note is that some of the graphs appear to show bimodal performance profiles (Permute, Storage, Snake, Towers, and Richards), or in the case of List, trimodal.

\section{Individual Annotations}
\begin{figure*}
	\input{graphs/pattern}
	\caption{Graphs of typing latices and performance with only an individual type annotation.}
	\label{f:pattern}		
\end{figure*}
Figures

We hypothesise that the tri/bi-modal graphs shown in figure~\ref{f:full}, are caused by individual type annotations: i.e. certain annotations have a significant effect on performance. In order to investigate this, we ran further experiments by testing configurations with each \emph{individual} type annotation enable, and all others removed, we then used this information to approximate and see if this appears to show a relation with the results across the typing lattice.

Firgure X shows the patterns we were able to identify, in particular...

\label{s-individual}
